A weakly structured stem for human origins in Africa

### 2022-02-18 paper.tex
fixed tex bug and added discussin of figure 3C

\begin{table}[t!]
    \centering
    \scriptsize
    \begin{tabular}{lllrrrr}
        \toprule
        \textbf{Likelihood} & \textbf{Label} & \textbf{Population Pair} &
            \textbf{Divergence} & \textbf{Migration rate} & \textbf{Migration} \\
        & & & 
            \textbf{Time (kya)} & \textbf{per generation} & \textbf{duration (ky)} \\
        \midrule
        \textbf{Continuous Model} & & & & & \\
        LL= -115,500 & a & Stem 1, Stem 2 & 1,163 & 6.43e-5 & 1,028 \\
        & b & Stem 2, Nama & NA & 5.82e-5 & 130 \\
        & c, d & Stem 2, Other Africans* & NA & 3.10e-5, \textbf{1.64e-4} & 130, 55 \\
        & e, f & Nama, Other Africans* & 135 & 4.1e-5, 9.8e-6 & 135, 60 \\
        & g & Mende, East Africans & 60 & \textbf{2.14e-4} & 60 \\
        & h & East Africans, British & 50 & 4.17e-5 & 50 \\
        & i & Gumuz, Amhara/Oromo & 12 &\textbf{3.36e-4} & 12 \\
        \textbf{Merger Model} & & & & & \\
        LL= -102,600 & a & Stem 1, Stem 2 & 1,442 & \textbf{1.16e-4} & 963 \\
        & -- & Stem 1S, Stem 1E & 479 & 0 (fixed) & -- \\
        & b & Stem 2 to Nama & 119 & \textbf{0.71} & pulse \\
        & c & Stem 2 to Stem 1E & 98 & \textbf{0.50} & pulse \\
        & d & Stem 2 to Mende & 25 & \textbf{0.18} & pulse \\
        & e, f & Nama, Other Africans* & 119 & 4.4e-5, 7.1e-6 & 119, 60 \\
        & g & Mende, East Africans* & 60 & \textbf{1.98e-4} & 60 \\
        & h & East Africans, British & 50 & 3.87e-5 & 50 \\
        & i & Gumuz, Amhara/Oromo &12 & \textbf{3.59e-4} & 12 \\
        \bottomrule
    \end{tabular}
    \caption{
        \textbf{Migration and divergence parameters from best fit models.}
        Labeled migration rates correspond to symmetric continuous migration
        bands shown in Figure~\ref{fig:2}. Both the continuous migration and
        merger models inferred a relatively deep split of human stem branches,
        though they were connected by ongoing migration that maintained their
        genetic similarity. Migration rates above \sgcomment{XX} are highlighted in bold}
        *: The ancestors of the Nama diverged from other African groups
        $\sim$120--135kya, which later branch into West and East African
        groups 60kya. Migration rates and durations are shown for 1) East Africans
        and their ancestors, and 2) the Mende, respectively.
    }
    \label{tab:migration-rates}
\end{table}

@@

Allowing for continuous migration between the stem populations substantially
improved the fits relative to zero migration ($LL \approx -102,600$ vs.
$-107,700$ in the merger model and $LL \approx -115,500$ vs. $-126,600$ with
continuous migration). With continuous migration between stems, population
structure extends back to 1.1--1.4Mya (Table~\ref{tab:migration-rates}).
Migration between the stems in these models is moderate, with a fraction of
migrant lineages each generation estimated as
$m=6.43\times10^{-5}-1.16\times10^{-4}$. For comparison, this is similar
to inferred migration rates between connected contemporary populations over the
past 50kya (Table~\ref{tab:migration-rates}). This ongoing (or at least,
periodic) gene flow qualitatively distinguishes these models from previously
proposed archaic admixture models (Figure~\ref{fig:supp-possible-models}C) as
the early branches remain closely related (Figure 3).

>>>>

\begin{table}[t!]
    \centering
    \scriptsize
    \begin{tabular}{lllrrrr}
        \toprule
        \textbf{Likelihood} & \textbf{Label} & \textbf{Population Pair} &
            \textbf{Divergence} & \textbf{Migration rate} & \textbf{Migration} \\
        & & & 
            \textbf{Time (kya)} & \textbf{per generation} & \textbf{duration (ky)} \\
        \midrule
        \textbf{Continuous Model} & & & & & \\
        LL= -115,500 & a & Stem 1, Stem 2 & 1,163 & 6.43e-5 & 1,028 \\
        & b & Stem 2, Nama & NA & 5.82e-5 & 130 \\
        & c, d & Stem 2, Other Africans* & NA & 3.10e-5, \textbf{1.64e-4} & 130, 55 \\
        & e, f & Nama, Other Africans* & 135 & 4.1e-5, 9.8e-6 & 135, 60 \\
        & g & Mende, East Africans & 60 & \textbf{2.14e-4} & 60 \\
        & h & East Africans, British & 50 & 4.17e-5 & 50 \\
        & i & Gumuz, Amhara/Oromo & 12 &\textbf{3.36e-4} & 12 \\
        \textbf{Merger Model} & & & & & \\
        LL= -102,600 & a & Stem 1, Stem 2 & 1,442 & \textbf{1.16e-4} & 963 \\
        & -- & Stem 1S, Stem 1E & 479 & 0 (fixed) & -- \\
        & b & Stem 2 to Nama & 119 & \textbf{0.71} & pulse \\
        & c & Stem 2 to Stem 1E & 98 & \textbf{0.50} & pulse \\
        & d & Stem 2 to Mende & 25 & \textbf{0.18} & pulse \\
        & e, f & Nama, Other Africans* & 119 & 4.4e-5, 7.1e-6 & 119, 60 \\
        & g & Mende, East Africans* & 60 & \textbf{1.98e-4} & 60 \\
        & h & East Africans, British & 50 & 3.87e-5 & 50 \\
        & i & Gumuz, Amhara/Oromo &12 & \textbf{3.59e-4} & 12 \\
        \bottomrule
    \end{tabular}
    \caption{
        \textbf{Migration and divergence parameters from best fit models.}
        Labeled migration rates correspond to symmetric continuous migration
        bands shown in Figure~\ref{fig:2}. Both the continuous migration and
        merger models inferred a relatively deep split of human stem branches,
        though they were connected by ongoing migration that maintained their
        genetic similarity. Bold text indicates migration rates above \sgcomment{XX} 
        *: The ancestors of the Nama diverged from other African groups
        $\sim$120--135kya, which later branch into West and East African
        groups 60kya. Migration rates and durations are shown for 1) East Africans
        and their ancestors, and 2) the Mende, respectively.
    }
    \label{tab:migration-rates}
\end{table}

@@

Allowing for continuous migration between the stem populations substantially
improved the fits relative to zero migration ($LL \approx -102,600$ vs.
$-107,700$ in the merger model and $LL \approx -115,500$ vs. $-126,600$ with
continuous migration). With continuous migration between stems, population
structure extends back to 1.1--1.4Mya (Table~\ref{tab:migration-rates}).
Migration between the stems in these models is moderate, with a fraction of
migrant lineages each generation estimated as
$m=6.43\times10^{-5}-1.16\times10^{-4}$. For comparison, this is similar
to inferred migration rates between connected contemporary populations over the
past 50kya (Table~\ref{tab:migration-rates}). This ongoing (or at least,
periodic) gene flow qualitatively distinguishes these models from previously
proposed archaic admixture models (Figure~\ref{fig:supp-possible-models}C) as
the early branches remain closely related (Figure 3). Because of this 
relatedness, only $1\%$ to $4\%$ of genetic differentiation among modern populations can be traced back
to this early population structure (Figure 3 and supplementary text)
